---
title: "SVM Prediction - Titanic Competition"
author: "Szymon Pawłowski"
date: "4 01 2021"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Kaggle Titanic Competition
It is one of the first challenges every ML beginner should dive in. In this competition the main goal is to predict which passengers survived the Titanic shipwreck using given data and creating a ML model. Here the SVM is used for prediction. In another file on GitHub "titanic_kaggle_competition.R" a whole work containing many approaches can be found.

## The data
The data has been split into two groups: training set and test set. Columns they contain:
* Survival - did passenger survive - 0=no, 1=yes
* Pclass - ticket class - 1=1st, 2=2nd, 3=3rd
* Name 
* Sex - sex - m/f 
* Age - age in years
* Sibsp - number of siblings / spouses aboard the Titanic 
* Parch - number of parents / children aboard the Titanic 
* Ticket - ticket number 
* Fare - passenger fare
* Cabin - cabin number 
* Embarked - port of embarkation - C=Cherbourg, Q=Quennstown, S=Southampton 

## Libraries


```{r libraries}
library(e1071)
library(ggplot2)
library(GGally)
library(fitdistrplus)
```

## Loading data

```{r loading}
df_t <- read.csv("test.csv")
df <- read.csv("train.csv")

```

## First look at data

```{r first look}
head(df)
tail(df)
summary(df)

df$Survived <- factor(df$Survived)
```
* Survival - need to transfer into factor
* Pclass - seems okay
* Name - could extract title from it
* Sex - need to transfer into numeric factor
* Age - contains missing values, need to replace them
* Sibsp and Parch - can get information of family size from here
* Ticket and Cabin - hard to get information so we will drop it for now
* Fare - seems okay, some NA's
* Embarked - need to transfer into numeric factor

# Exploratory Data Analysis (EDA)

## Pclass
```{r edap, echo=FALSE}
ggplot(df, aes(x=Pclass, fill=Survived)) + geom_histogram(stat="count")
```
We see that in the 1st class there were more survivors and in the 3rd class the number of non-survivors is relatively high comparing to the number of survivors.


## Sex
```{r edas, echo=FALSE}
ggplot(df, aes(x=Sex, fill=Survived)) + geom_histogram(stat="count")
```
Most of the survivors were females. This is a well-known fact that women with children are in the first place to be saved.

## Age
```{r edaa, echo=FALSE}
ggplot(df, aes(x=Age, fill=Survived)) + geom_density(alpha=0.5)
```
Because of the low correlation it is hard to tell something valuable. Most survivors were between age of 20 and 30, similarly the non-survivrors. Later we will try to categorize age and check out if we can get more informations.

## Fare
```{r edaf, echo=FALSE}
ggplot(df, aes(x=Fare, fill=Survived)) + geom_density(alpha=0.5)
```
The less is the charge the less survivors. It might be caused with a fact that well paid cabins could be more durable or be placed closer to rescue boats.

## Embarked
```{r edae, echo=FALSE}
ggplot(df, aes(x=Embarked, fill=Survived)) + geom_histogram(stat="count")
```
The are more survivors in Cherbourg but less in Queenstown or Southampton. Let's see if there is some relationship with Fares.
```{r edae2, echo=FALSE}
ggplot(df,aes(x=Fare,fill=factor(Embarked))) + geom_density(alpha=0.5)
```
As we can see in Queenstown or Southampton most Fares are in the low interval while in the Cherbourg we may notice that Fares were aswell low as high so the reason standing behind more survivors in Cherbourg are probably better paid cabins.
# Feature Engineering 

## Age
Allocating median instead of NAs.
```{r fage}
df[is.na(df$Age)==TRUE, ]$Age <- median(df[is.na(df$Age)==FALSE, ]$Age)
df_t[is.na(df_t$Age)==TRUE, ]$Age <- median(df_t[is.na(df_t$Age)==FALSE, ]$Age)

df["AgeBin"]<-NA
df[df$Age < 18,]$AgeBin <- (-1) #kid
df[df$Age >=18 & df$Age<50,]$AgeBin <- 0 #adult
df[df$Age>=50,]$AgeBin <- 1 #elderly

df_t["AgeBin"]<-NA
df_t[df_t$Age < 18,]$AgeBin <- (-1) #kid
df_t[df_t$Age >=18 & df_t$Age<50,]$AgeBin <- 0 #adult
df_t[df_t$Age>=50,]$AgeBin <- 1 #elderly
```

## Sex
```{r fsex}
df[df$Sex=="male","Sex"]<-(-1)
df[df$Sex=="female","Sex"]<-1
df$Sex<-as.numeric(df$Sex)

df_t[df_t$Sex=="male","Sex"]<-(-1)
df_t[df_t$Sex=="female","Sex"]<-1
df_t$Sex<-as.numeric(df_t$Sex)
```

## Ticket and Cabin

```{r ftc}
df$Cabin<-NULL
df$Ticket<-NULL

df_t$Cabin<-NULL
df_t$Ticket<-NULL
```

## Name

```{r fn}
df["Title"]<-sub("\\s.*","",sub(".*,\\s","",df$Name))
df["Title"]<-factor(df$Title)
meaning<-unique(df$Title)
df$Title<-as.numeric(df$Title)
numeric<-unique(df$Title)
meaning <- data.frame(numeric=numeric, meaning=meaning)
t(meaning)
df[df$Title!=9 & df$Title!=12 & df$Title!=13 & df$Title!=14,]$Title <-0 #Other
df[df$Title==12 | df$Title==13,]$Title <- (-1) #Mr, #Mrs
df[df$Title==9 | df$Title==14,]$Title <- 1 #Miss/Ms


df_t["Title"]<-sub("\\s.*","",sub(".*,\\s","",df_t$Name))
df_t["Title"]<-factor(df_t$Title)
meaning_t<-unique(df_t$Title)
df_t$Title<-as.numeric(df_t$Title)
numeric_t<-unique(df_t$Title)
meaning_t <- data.frame(numeric=numeric_t, meaning=meaning_t)
t(meaning_t)
df_t[df_t$Title!=6 & df_t$Title!=7& df_t$Title!=5& df_t$Title!=8,]$Title <-0
df_t[df_t$Title==6 | df_t$Title==7,]$Title <- (-1)
df_t[df_t$Title==5 | df_t$Title==8,]$Title <- 1

df$Name <- NULL
df_t$Name <- NULL

```
```{r edat, echo=FALSE}
ggplot(df, aes(x=Title, fill=Survived)) + geom_histogram(stat="count")
```
We see that most of the non-survivors were title Mr. while survivors - Ms. or Mrs..

## Embarked
```{r fe}
df[df$Embarked=="C",]$Embarked <- (-1)
df[df$Embarked=="Q",]$Embarked <- 0
df[df$Embarked=="S",]$Embarked <- 1
df$Embarked<-as.numeric(df$Embarked)

df_t[df_t$Embarked=="C",]$Embarked <- (-1)
df_t[df_t$Embarked=="Q",]$Embarked <- 0
df_t[df_t$Embarked=="S",]$Embarked <- 1
df_t$Embarked<-as.numeric(df_t$Embarked)
summary(df)
```

We got 2 NA's, will replace them with random number
```{r fe2}

df[is.na(df$Embarked)==TRUE,]$Embarked <- sample(c(1,2,3),2)
summary(df)
```

## Fare
Getting rid off NA in Fare test data
```{r ff}
summary(df_t)
fare <- df_t[is.na(df_t$Fare)==FALSE,]$Fare
df_t[is.na(df_t$Fare)==TRUE,]$Fare<-rnorm(1,mean(fare),sd(fare))
```

## Family Size
```{r ffs}
df["Fsize"]<-df$Parch + df$SibSp + 1
df_t["Fsize"] <- df_t$Parch + df_t$SibSp +1
```
```{r edafs, echo=FALSE}
ggplot(df, aes(x=Fsize, fill=Survived)) + geom_histogram(stat="count")
```
In the families of size 2-4 there were more survivors and in the families of size 1 or 5 and more there were more non-survivors. Most families of size 2-4 consists of married couples and children and as we said, they were a priority to be saved.

```{r ffs2}
df[df$Fsize==1,]$Fsize <- (-1) #single
df[df$Fsize>1 & df$Fsize<5,]$Fsize <- 0 #family
df[df$Fsize>=5,]$Fsize <- 1 #large family

df_t[df_t$Fsize==1,]$Fsize <- (-1) #single
df_t[df_t$Fsize>1 & df_t$Fsize<5,]$Fsize <- 0 #family
df_t[df_t$Fsize>=5,]$Fsize <- 1 #large family

```

```{r edafs2, echo=FALSE}
ggplot(df, aes(x=Fsize, fill=Survived)) + geom_histogram(stat="count")
```

Creating correlogram to see relationships between features.
```{r eda1, echo=FALSE}
df$Survived<-as.factor(df$Survived)
ggpairs(df,columns=c(1,3:11),ggplot2::aes(colour=Survived))
```
# Creating SVM model
Now we will perform SVM using radial kernel and choosing best parameters with tune().
```{r model1}
tune.out<-tune(svm, Survived~Pclass+Sex+Fare+Embarked+Title+Age+Fsize, data=df,
               kernel="radial",type="C-classification", 
               range = list(cost=10^(-2:2),gamma=10^(-5:5)))

summary(tune.out)
svm_fit<-tune.out$best.model
```
Checking training error rates.
```{r model2}
table(svm_fit$fitted,df$Survived)
```
Making final prediction and saving result into data frame.
```{r model3}
prediction=predict(svm_fit,df_t)

svm_p<-data.frame(PassengerId=892:1309, Survived=prediction) 
```
Writing into .csv file ready for a submission on Kaggle.
```{r model4}
write.csv(svm_p,file="svm_prediction.csv", row.names=F)
```

